ShrinkNets

### 2018-01-06 paper/paper.tex
Fix technical innacuracies

When designing neural networks, one of the key parameters is the network size, i.e., 
%, number of layers
%and number of neuros per layer, 
%(width and depth)
the number of layers and neurons per layer.  Choosing these parameters appropriately
can dramatically affect performance, yet
%In particular, these hyper parameters have a strong correlation with
%over/underfitting \srm{give reference} \gl{I did some litterature review and
%actually this statement is quite controversial, should we rephrase ?}. 
there is no reliable way to efficiently set them. Although many search
 strategies and heuristics~\cite{Bengio2012a} have been proposed, including
%To select the best candidates, researchers have used strategies such
%as 
random search,
%\gl{Should we also move the references here if we remove the
%pargraph from the related work ?}, 
meta-gradient descent~\cite{Pedregosa2016},
Gaussian processes~\cite{Bergstra2011a}, and Parzen Estimators~\cite{Bergstra2011a}. 
%to find good parameters without exhaustively
%exploring the entire hyper-parameter space. 
 these methods generally require a compute-intensive
search of parameter space.

@@

In this paper we present a method to automatically find an appropriate network
size,
%from a single parameter, 
drastically reducing optimization time.
%dimensionality. 
The key idea is to
\emph{learn} the right network size at the same time that the network is
learning the main task. For example, for an image classification task, with our
approach we can provide the training data to a network---without sizing it a
priori---and expect to end up with a network that has learned to classify images
with an accuracy similar to a the best manually engineered network.
%\gl{too strong, we do not really have
%  any guarantee on the generalization, how about changing it to: to end up with
%  a network that learnt a tradeoff between size and accuracy}. 
Our approach has two main benefits. First, no longer need to choose a
network size before training. Second, the final network size will tuned to be
appropriate for the task at hand, and not larger. 
%Because the sizing occurs
%dynamically, while the
%network is training, we remove the useless neurons 
%implement a new mechanism that removes
%the deactivated neurons to  
%\gl{Should we talk about the
%fact that it automatically gets rid of dead neurons ?}. 
This is important
because over-sized networks have a lower inference throughput and higher memory
footprint.

@@

\textbf{Filter Layers: } Filter layers have weights in the range $[0,+\infty]$ and are placed after
linear and convolutional layers.
%\gl{This is how I
%  used it but with the new implementation we can be more creative and put them
%  anywhere, should we say "usually placed" insdead ?}. 
The \textit{Filter Layer} takes an input of size $\left(B \times C \times D_1
  \times \dots \times D_n\right)$, where $B$ is the batch size, $C$ the number
of features (or channels, in the case of convolutional layers), and $D$ any
additional dimension. This structure makes it compatible with fully connected
layers with $n=0$ or convolutional layers with $n=2$. Their crucial property is
a parameter $\theta \in \mathbb{R}^C$, defined as follows: \vspace{-1em}
\begin{equation} Filter(I;\theta) = I \circ \max(0, \theta) \end{equation}
Where $\circ$ is the pointwise multiplication, and $\theta$ is expanded in all
dimensions to match the input size (except the second one since they are equal by definition). It is easy to see that if for any
$k$, if $\theta_k \leq 0$, the $k^{\text{th}}$ feature/channel will forever be
$0$. When this happens, we say the Filter layer deactivates the neuron.
\srm{Not easy to see this because you haven't told me that theta only decays
  over time?} \gl{The decay is only caused by the loss, and actually it does not depends on how it became zero, if it is zero for any reason, multiplying by zero will nullfy any value and cancel any gradient going through the layer} These disabled neurons/channels can be removed from the network
without changing its output. Before explaining how that is achieved, we explain
next how the weights of the Filter Layer are initialized and adjusted during training.

@@

%\subsection{The training procedure}
\textbf{Training Procedure: } Once Filter layers are placed in a network and initialized (sampled from the Uniform$[0, 1]$ distribution),
%To train networks we need start with a substantially oversized network, then we
%insert \textit{Filter Layers}  (usually after every linear or convolutional
%layer except the last one) and we sample their weight from the
%$\text{Uniform}(0, 1)$ distribution. 
we could train the network directly using our standard loss function, and we could achieve performance equivalent to a normal neural
network. However, our goal is to find the smallest network with reasonable
performance. We achieve that by introducing sparsity in the parameters of the
\textit{Filter Layers}, thus forcing the deactivation of neurons \srm{by decaying the value of theta} \gl{$\theta$ only decays (and only for some features), only because of the loss function that we have not defined yet at this moment in the paper}
%. Indeed, having a negative component in the $\theta$
%parameter of the filter layer permamently disable its associated feature
%\gl{Maybe redundant ? we talked about that in the previous paragraph} . 
To obtain this sparsity, we simply redefine the loss function:
\vspace{-.5em}
\begin{equation}
  L'(x,y;\theta) = L(x, y) + \lambda|\theta|
\end{equation}

The additional term $\lambda|\theta|$ introduces sparsity (see Lasso
loss~\cite{Tibshirani1996}). 
% The $\lambda$ parameter, that can take any
% positive value, adjusts how aggressively the network deactivates neurons, with
% larger values indicating more aggressive deactivation.
Indeed, the second component of the loss, because it increases the gradient in
respect with $\theta$, pushes its value towards zero. Neurons with little
impact on the original loss (gradient lower than $\lambda$), will not be able
to compete against this attraction towards zero. Because the entries in
$\theta$ with a value of $0$ correspond to dead neurons, $\lambda$ effectively
controls the number of neurons/channels in the entire network.  Next, we
explain how to implement ShrinkNets efficiently.

@@

In this experiment we want to find an appropriate network size for two network
architectures, a multi-layer perceptron (MLP) and the \textit{LeNet-5}
model~\cite{Lecun1998}. We have no prior information about the network size, but
we set an upper bound of $50$ channels per convolutional layer and $5000$
neurons for fully connected layers. We then use random search to vary the
parameters that determine the network size, i.e., $\lambda$ in the case of
ShrinkNets, random selection of depth and witdh in the case of \textsf{Fixed}
and similar to previous but adding a $L_2$ penalty~\cite{Ng2004}
(\textsf{Fixed-L2}).  We train all variants on \texttt{MNIST}~\cite{Lecun1998},
\texttt{FashionMNIST}~\cite{Xiao2017} and
\texttt{CIFAR10}~\cite{Krizhevsky2009}. We use the same number of epochs, 100
for MLPs and 200 for \textit{LeNet-5}. We select the epoch that performed the
best on the validation set and evaluated it on the testing set. We repeat this
process 50 times and show the distribution of the testing accuracy in
\autoref{hyper_opt_res}.

The results in the figure show that ShrinkNets finds better networks with fewer
training iterations than the other methods, because it achieves higher median 
and less variable accuracies. Despite the simple nature of this experiment, we
believe that the results generalize to more complex search methods.

@@

There are many methods in the literature that aim to simplify network
structure. Most of them focuses on removing connections (eg: ~\cite{Cun},
~\cite{Han2015}). On the other side, ShrinkNets and some others
~\cite{Scardapane2017} and ~\cite{Philipp} try to remove entire neurons instead
of connections. This is very useful because it reduces the size of the matrices,
producing speed-up on every device. ShrinkNets improves on
~\cite{Scardapane2017} because it removes neurons during training, speeding up
the rest of the process. Further it outperforms~\cite{Philipp} on convergence
speed and because it does not need to change the optimizer. To the best of our
knowledge this is also the first work that tries to learn the number of channels
in a convolutional neural network.

>>>>

When designing neural networks, one of the key parameters is the network size, i.e., 
%, number of layers
%and number of neuros per layer, 
%(width and depth)
the number of layers and neurons per layer.  Choosing these parameters appropriately
can dramatically affect performance, yet
%In particular, these hyper parameters have a strong correlation with
%over/underfitting \srm{give reference} \gl{I did some litterature review and
%actually this statement is quite controversial, should we rephrase ?}. 
there is no reliable way to efficiently set them. Although many search
 strategies and heuristics~\cite{Bengio2012a} have been proposed, including
%To select the best candidates, researchers have used strategies such
%as 
random search,
%\gl{Should we also move the references here if we remove the
%pargraph from the related work ?}, 
meta-gradient descent~\cite{Pedregosa2016},
Gaussian processes~\cite{Bergstra2011a}, and Parzen Estimators~\cite{Bergstra2011a}, 
%to find good parameters without exhaustively
%exploring the entire hyper-parameter space. 
 they generally require a compute-intensive
search of parameter space.

@@

In this paper we present a method to automatically find an appropriate network
size,
%from a single parameter, 
drastically reducing optimization time.
%dimensionality. 
The key idea is to
\emph{learn} the right network size at the same time that the network is
learning the main task. For example, for an image classification task, with our
approach we can provide the training data to a network---without sizing it a
priori---and expect to end up with a network that has learned to classify images
with an accuracy similar to a the best manually engineered network.
%\gl{too strong, we do not really have
%  any guarantee on the generalization, how about changing it to: to end up with
%  a network that learnt a tradeoff between size and accuracy}. 
Our approach has two main benefits. First, we no longer need to choose a
network size before training. Second, the final network size will tuned to be
appropriate for the task at hand, and not larger. 
%Because the sizing occurs
%dynamically, while the
%network is training, we remove the useless neurons 
%implement a new mechanism that removes
%the deactivated neurons to  
%\gl{Should we talk about the
%fact that it automatically gets rid of dead neurons ?}. 
This is important
because over-sized networks have a lower inference throughput and higher memory
footprint.

@@

\textbf{Filter Layers: } Filter layers have weights in the range $[0,+\infty]$ and are usually placed after
linear and convolutional layers.
%\gl{This is how I
%  used it but with the new implementation we can be more creative and put them
%  anywhere, should we say "usually placed" insdead ?}. 
The \textit{Filter Layer} takes an input of size $\left(B \times C \times D_1
  \times \dots \times D_n\right)$, where $B$ is the batch size, $C$ the number
of features (or channels, in the case of convolutional layers), and $D$ any
additional dimension. This structure makes it compatible with fully connected
layers with $n=0$ or convolutional layers with $n=2$. Their crucial property is
a parameter $\theta \in \mathbb{R}^C$. The output is defined as follows: \vspace{-1em}
\begin{equation} Filter(I;\theta) = I \circ \max(0, \theta) \end{equation}
Where $\circ$ is the pointwise multiplication, and $\theta$ is expanded in all
dimensions to match the input size (except the second one since they are equal
by definition). It is easy to see that if for any $k$, if $\theta_k \leq 0$,
the $k^{\text{th}}$ input feature/channel is multiplied by zero and have no
influence on the output. If this happens, we say the Filter layer deactivates
the neuron. These disabled neurons/channels can be removed from the network
without changing its output. Before explaining how that is achieved, we explain
next how the weights of the Filter Layer are initialized and adjusted during
training.

@@

%\subsection{The training procedure}
\textbf{Training Procedure: } Once Filter layers are placed in a network and initialized (sampled from the Uniform$[0, 1]$ distribution),
%To train networks we need start with a substantially oversized network, then we
%insert \textit{Filter Layers}  (usually after every linear or convolutional
%layer except the last one) and we sample their weight from the
%$\text{Uniform}(0, 1)$ distribution. 
we could train the network directly using our standard loss function, and we could achieve performance equivalent to a normal neural
network. However, our goal is to find the smallest network with reasonable
performance. We achieve that by introducing sparsity in the parameters of the
\textit{Filter Layers}, thus forcing the deactivation of neurons%. Indeed, having a negative component in the $\theta$
%parameter of the filter layer permamently disable its associated feature
%\gl{Maybe redundant ? we talked about that in the previous paragraph} . 
. To obtain this sparsity, we simply redefine the loss function:
\vspace{-.5em}
\begin{equation}
  L'(x,y;\theta) = L(x, y) + \lambda|\max(0, \theta)|
\end{equation}

The additional term $\lambda|\max(0, \theta)|$ introduces sparsity (see Lasso
loss~\cite{Tibshirani1996}). 
% The $\lambda$ parameter, that can take any
% positive value, adjusts how aggressively the network deactivates neurons, with
% larger values indicating more aggressive deactivation.
The second component of the loss, because it increases the gradient with
respect to $\theta$, pushes its value towards zero. Neurons with little impact
on the original loss (gradient lower than $\lambda$), will not be able to
compete against this attraction towards zero. Because the entries in $\theta$
with a value of $0$ or less correspond to dead neurons, $\lambda$ effectively
controls the number of neurons/channels in the entire network. We introduced
the $\max(\dots)$ into the loss to make sure that neurons are permamently disabled
when performing gradient descent based optimization. Next, we
explain how to implement ShrinkNets efficiently.

@@

In this experiment we want to find an appropriate network size for two network
architectures, a multi-layer perceptron (MLP) and the \textit{LeNet-5}
model~\cite{Lecun1998}. We have no prior information about the network size,
but we set an upper bound of $50$ channels per convolutional layer and $5000$
neurons for fully connected layers. We then use random search to vary the
parameters that determine the network size, i.e., $\lambda$ in the case of
ShrinkNets, the width of each layer in the case of
\textsf{Fixed} and similar to previous but adding a $L_2$ penalty~\cite{Ng2004}
(\textsf{Fixed-L2}).  We train all variants on \texttt{MNIST}~\cite{Lecun1998},
\texttt{FashionMNIST}~\cite{Xiao2017} and
\texttt{CIFAR10}~\cite{Krizhevsky2009}. We use the same number of epochs, 100
for MLPs and 200 for \textit{LeNet-5}. We select the epoch that performed the
best on the validation set and evaluated it on the testing set. We repeat this
process 50 times and show the distribution of the testing accuracy in
\autoref{hyper_opt_res}.

The results in the figure show that ShrinkNets finds better networks with fewer
training iterations than the other methods, because it achieves higher median 
and less variable accuracies. Despite the simple nature of this experiment, we
believe that the results generalize to more complex search methods; mainly
because our solution reduces the dimensionality of the space they have to
explore.

@@

There are many methods in the literature that aim to simplify network
structure. Most of them focuses on removing connections (eg: ~\cite{Cun},
~\cite{Han2015}). On the other side, ShrinkNets and some others
~\cite{Scardapane2017} and ~\cite{Philipp} try to remove entire neurons instead
of connections. This is very useful because it reduces the size of the matrices,
producing speed-up even on devices/libraries that only support dense matrices
. ShrinkNets improves on
~\cite{Scardapane2017} because it removes neurons during training, speeding up
the rest of the process. Further it outperforms~\cite{Philipp} on convergence
speed and because it does not need to change the optimizer. To the best of our
knowledge this is also the first work that tries to learn the number of channels
in a convolutional neural network.
